---
title: "Class Project"
author: "Hukai Luo"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=6, fig.height=3.4,echo=TRUE, warning=FALSE, message=FALSE)
```

# 1 Simulate Geometric Brownian Motion 

Geometric Brownian Motion solution \[S(T)=S(t)e^{(\mu-\frac{1}{2}\sigma^2)(T-t)+\sigma\sqrt{T-t}z}\]
Now, given some initial data $S(0)=100$, $r=0.04$, $\sigma=0.3$, $T=1$, let's simulate the Geometric Brownian Motion pathways.
```{r code1,fig.height=3}
s0 <- 100
r <- 0.04
sigma <- 0.3                                 # input initial data
T <- 1
S_T <- function(s0,r,sigma,T,n=19){          # build the GBM function below
  data <- double(0)
  data[1] <- s0
  for(i in 1:19){
    s0 <- s0*exp((r-0.5*sigma^2)*(T/n)+sigma*sqrt(T/n)*rnorm(1,0,1))
    data[i+1] <- s0
  }
  return(data)
}
data <- S_T(s0,r,sigma,T)                    # use the function the get each path's stock value S_T
library(ggplot2)                             # plot the pathway
plot1 <- ggplot(data.frame(x=seq(1:20),S_t = data),aes(x=x,y=S_t))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="GBM Pathway", subtitle="T=1,sigma=0.3,S(0)=100,r=0.04,n=19")
plot1
```


```{r multiplot,echo=FALSE}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# 2 Vanilla black-scholes european call option

Set the $K= 100$
```{r code2}
K=100                                 # Excercise price K
n <- 19
B_S <- function(data,K){              # Build the Bs call option function
  call_value <- double(0)             # calculate d1,d2,call option value,cashflow,replicating value
  d1 <- double(0)
  d2 <- double(0) 
  for(i in 1:20){
    d1[i] <- (log(data[i]/K)+(r+0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    d2[i] <- (log(data[i]/K)+(r-0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    call_value[i] <- data[i]*pnorm(d1[i])-K*exp(-r*(20-i)*T/n)*pnorm(d2[i])
    }
  cashflow <- double(0)               # calculate the cash flow and replicating value
  bt <- double(0)
  replicate <- double(0)
  bt[1] <- K*exp(-r*T)*pnorm(d2[1])
  cashflow[1] <- 0
  replicate[1] <- -bt[1]+data[1]*pnorm(d1[1])
  for(i in 2:20){
    cashflow[i] <- data[i]*(pnorm(d1[i])-pnorm(d1[i-1]))
    bt[i] <- bt[i-1]*exp(r*T/n)+cashflow[i]
    replicate[i] <- data[i]*pnorm(d1[i])- bt[i]
  }
  return(data.frame(stock=data,d1=d1,delta=pnorm(d1),Bt=bt,replicate=replicate,call_option=call_value))
}
call <- B_S(data,100)
print(call)
```
```{r plot2,fig.height=7,echo=FALSE}
plot2 <- ggplot(data.frame(x=seq(1:20),call_value = call$call_option),aes(x=x,y=call_value))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="BS call option value", subtitle="K=100")
plot3 <- ggplot(data.frame(x=seq(1:20),delta = call$replicate),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Replicating portfolio value", subtitle="K=100")
plot5 <- ggplot(data.frame(x=seq(1:20),delta = call$stock),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Underlying stock value", subtitle="K=100")

layout <- matrix(c(1,2,3),3,1,byrow=TRUE)
multiplot(plot5,plot2, plot3,layout=layout)
```

The value of the replicating portfolio at time T \[C=\triangle S+B\]
The terminal value of the call at time T \[C=max(S-K,0)\]
Now repeat above process for 2000 times to generate the distribution of replicating error.
```{r code3}
stock <- call <- replicate <- error <- double(0)
for(i in 1:2000){
  t <- B_S(S_T(s0,r,sigma,T),100)
  error[i] <- t$replicate[20]-t$call_option[20]
  replicate[i] <- t$replicate[20]
  call[i] <- t$call_option[20]
  stock[i] <- t$stock[20]
}
```
```{r 12,fig.height=6,fig.width=6,echo=FALSE}
plot4<- ggplot(data.frame(x = error), aes(x = x))+ 
  geom_histogram(aes(y=..density..),fill="steelblue")+labs(x = 'r(t)', 
  y = 'density', title='Error distribution density')
plot13 <- ggplot(data.frame(x=stock,y1=replicate,y2=call))+
  geom_point(aes(x=x,y=y1),size=0.5,color='steelblue')+
  geom_point(aes(x=x,y=y2),size=0.5,color='orange')+
  labs(title = "Value distribution",subtitle = "Replicate/BS call option",y="Option Value",x="Stock Value")
layout <- matrix(c(1,2),2,1,byrow=TRUE)
multiplot(plot4,plot13,layout=layout)
```


# 3 Cash-or-nothing call option

Cash-or-nothing call option either pays you a fixed amount of money or nothing at all.\[C_{cn}=ke^{-r\tau}N(d2)\]
where $S$ the initial stock price, $K$ the strike price, $T$ the time to maturity,$\sigma$ the volatility and r the risk free interest rate, in thie section, we will set $k=1$ for convenience.
```{r code4,fig.height=3}
t <- 1e-6
C_N <- function(data,K){
  call_value <- double(0)             # calculate d1,d2,call option value,cashflow,replicating value
  d1 <- double(0)
  d2 <- double(0)
  delta <- double(0)
  for(i in 1:20){
    d1[i] <- (log(data[i]/K)+(r+0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    d2[i] <- (log(data[i]/K)+(r-0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    call_value[i] <- exp(-r*(20-i)*T/n)*pnorm(d2[i])
    delta[i] <- exp(-r*(20-i)*T/n)*dnorm(d2[i])/(data[i]*sigma*sqrt((20-i)*T/n))
  }
  delta[20] <- 0
  cashflow <- double(0)               # calculate the cash flow and replicating value
  bt <- double(0)
  replicate <- double(0)
  bt[1] <- -call_value[1]+data[1]*delta[1]
  cashflow[1] <- 0
  replicate[1] <- -bt[1]+data[1]*delta[1]
  for(i in 2:20){
    cashflow[i] <- data[i]*(delta[i]-delta[i-1])
    bt[i] <- bt[i-1]*exp(r*T/n)+cashflow[i]
    replicate[i] <- data[i]*delta[i]- bt[i]
  }
  return(data.frame(stock=data,delta=delta,bt=bt,replicate=replicate,CON=call_value))
}
call1 <- C_N(data,100)
print(call1)
```

```{r plot3,fig.height=7,echo=FALSE}
plot6 <- ggplot(data.frame(x=seq(1:20),call_value = call1$CON),aes(x=x,y=call_value))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Cash-or-nothing option value", subtitle="K=100")
plot7 <- ggplot(data.frame(x=seq(1:20),delta = call1$replicate),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Replicating portfolio value", subtitle="K=100")
plot8 <- ggplot(data.frame(x=seq(1:20),delta = call1$stock),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Underlying stock value", subtitle="K=100")

layout <- matrix(c(1,2,3),3,1,byrow=TRUE)
multiplot(plot8,plot6, plot7,layout=layout)
```

The value of the replicating portfolio at time T \[C=\triangle S+B\]
The terminal value of the call at time T \[C=\frac{max(S-K,0)}{|S-K|}\]
Now repeat above process for 2000 times to generate the distribution of replicating error.
```{r code5}
stock <- call <- replicate <- error <- double(0)
for(i in 1:2000){
  t <- C_N(S_T(s0,r,sigma,T),100)
  error[i] <- t$replicate[20]-t$CON[20]
  replicate[i] <- t$replicate[20]
  call[i] <- t$CON[20]
  stock[i] <- t$stock[20]
}
```
```{r 13,fig.height=5.5,fig.width=6,echo=FALSE}
plot9<- ggplot(data.frame(x = error), aes(x = x))+ 
  geom_histogram(aes(y=..density..),fill="steelblue")+labs(x = 'r(t)', 
  y = 'density', title='Error distribution density')
plot14 <- ggplot(data.frame(x=stock,y1=replicate,y2=call))+
  geom_point(aes(x=x,y=y1),size=0.5,color='steelblue')+
  geom_point(aes(x=x,y=y2),size=0.5,color='orange')+
  labs(title = "Value distribution",subtitle = "Replicate/CON call option",y="Option Value",x="Stock Value")
layout <- matrix(c(1,2),2,1,byrow=TRUE)
multiplot(plot9,plot14,layout=layout)
```

# 4 Asset-or-nothing call option

The asset-or-nothing option is basically the same, but your payment equals the price of the asset underlying the option.\[C_{an}=SN(d1)\]
In order to replicate this option, we can buy call at $K=100$ and buy K units cash-or-nothing call at $K=100$.

```{r code6,fig.height=3}
epsilon <- 0.01
A_N <- function(data,K){
  call_value <- double(0)             # calculate d1,d2,call option value,cashflow,replicating value
  d1 <- double(0)
  d2 <- double(0)
  delta <- double(0)
  for(i in 1:20){
    d1[i] <- (log(data[i]/K)+(r+0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    d2[i] <- (log(data[i]/K)+(r-0.5*sigma^2)*(20-i)*T/n)/(sigma*sqrt((20-i)*T/n))
    call_value[i] <- data[i]*pnorm(d1[i])
    delta[i] <- pnorm(d1[i])+dnorm(d1[i])/(sigma*sqrt((20-i)*T/n))
  }
  delta[20] <- pnorm(d1[i])
  cashflow <- double(0)               # calculate the cash flow and replicating value
  bt <- double(0)
  replicate <- double(0)
  bt[1] <- -call_value[1]+data[1]*delta[1]
  cashflow[1] <- 0
  replicate[1] <- -bt[1]+data[1]*delta[1]
  for(i in 2:20){
    cashflow[i] <- data[i]*(delta[i]-delta[i-1])
    bt[i] <- bt[i-1]*exp(r*T/n)+cashflow[i]
    replicate[i] <- data[i]*delta[i]- bt[i]
  }
  return(data.frame(stock=data,delta=delta,bt=bt,replicate=replicate,AON=call_value))
}
call2 <- A_N(data,100)
print(call2)
```

```{r plot4,fig.height=7,echo=FALSE}
plot9 <- ggplot(data.frame(x=seq(1:20),call_value = call2$AON),aes(x=x,y=call_value))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Asset-or-nothing option value", subtitle="K=100")
plot10 <- ggplot(data.frame(x=seq(1:20),delta = call2$replicate),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Replicating portfolio value", subtitle="K=100")
plot11 <- ggplot(data.frame(x=seq(1:20),delta = call2$stock),aes(x=x,y=delta))+
  geom_line(col="steelblue", size=0.7)+scale_x_continuous(breaks=seq(1, 20, 1))+
  labs(title="Underlying stock value", subtitle="K=100")

layout <- matrix(c(1,2,3),3,1,byrow=TRUE)
multiplot(plot11,plot9, plot10,layout=layout)
```

The value of the replicating portfolio at time T \[C=\triangle S+B\]
The terminal value of the call at time T \[C=\frac{max(S-K,0)}{|S-K|}S\]
Now repeat above process for 2000 times to generate the distribution of replicating error.
```{r code7}
stock <- call <- replicate <- error <- double(0)
for(i in 1:2000){
  t <- A_N(S_T(s0,r,sigma,T),100)
  error[i] <- t$replicate[20]-t$AON[20]
  replicate[i] <- t$replicate[20]
  call[i] <- t$AON[20]
  stock[i] <- t$stock[20]
}
```
```{r 14,fig.height=5.5,fig.width=6,echo=FALSE}
plot12<- ggplot(data.frame(x = error), aes(x = x))+ 
  geom_histogram(aes(y=..density..),fill="steelblue")+labs(x = 'r(t)', 
  y = 'density', title='Error distribution density')
plot15 <- ggplot(data.frame(x=stock,y1=replicate,y2=call))+
  geom_point(aes(x=x,y=y1),size=0.5,color='steelblue')+
  geom_point(aes(x=x,y=y2),size=0.5,color='orange')+
  labs(title = "Value distribution",subtitle = "Replicate/AON call option",y="Option Value",x="Stock Value")
layout <- matrix(c(1,2),2,1,byrow=TRUE)
multiplot(plot12,plot15,layout=layout)
```





















